---
title: "Data Analysis Notebook"
output: html_notebook
---

```{r}

#Authors: [Sydney Stanton, Carter Zborowski]

#Author Date: 06/09/2025
#Purpose: The purpose of this notebook is to house all data set transformation, cleansing, visualization, statistical analysis, and note-taking for the 2025 CU Athletic Department Sports Science Internship Program

#LAST UPDATED: 06/12/2025

#Including helpful libraries
library(tidyverse)
library(readxl)
library(aod)
library(tidyverse)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lubridate)
```

```{r}
#load all of the useful datasets
ACWR <- read.csv("~/T2P1-data-analysis-2025/data-sets/ACWR-Catapult FB.csv")
catapult <- read.csv("~/T2P1-data-analysis-2025/data-sets/Catapult Session - Outdoor FB.csv")
incident <- read.csv("~/T2P1-data-analysis-2025/data-sets/Incident Report FB.csv")
perfnorm <- read.csv("~/T2P1-data-analysis-2025/data-sets/Performance Normative Data FB.csv")
perfrisk <- read.csv("~/T2P1-data-analysis-2025/data-sets/Performance Risk Assessment FB.csv")
riskstatus <- read.csv("~/T2P1-data-analysis-2025/data-sets/Soft Tissue Risk Status Report FB.csv")
strengthtest <- read.csv("~/T2P1-data-analysis-2025/data-sets/Strength Testing Data FB.csv")
VALD <- read.csv("~/T2P1-data-analysis-2025/data-sets/VALD - Performance Test FB.csv")
wellness <- read.csv("~/T2P1-data-analysis-2025/data-sets/Wellness Report FB.csv")
```

## Data Cleaning 

1.ACWR- Catapult 
```{r}
ACWR <- read.csv("~/T2P1-data-analysis-2025/data-sets/ACWR-Catapult FB.csv")
```

```{r}
#refined ACWR dataset 
ACWR <- ACWR %>%
  filter(Sport=="FOOTBALL")
Acwr <- subset(ACWR,select=c("anon_id","Date","Position","Total.Player.Load","Acute.Total.Acceleration.Load.EWMA","Acceleration.Load.ACWR","Player.Load.ACWR","Total.Distance.ACWR","Month","Year"))

#cut back to last 12 months (last year)
Acwr$Date <- mdy(Acwr$Date)
current_date <- Sys.Date()
start_date <- current_date- months(12)

filtered_Acwr <- Acwr %>%
  filter(Date >= start_date) %>%
  group_by(anon_id,Date,Month,Year,Position) %>%
  summarise(Total.Player.Load=mean(Total.Player.Load),
            Total.Distance.ACWR=mean(Total.Distance.ACWR),
            Player.Load.ACWR= mean(Player.Load.ACWR),
            Acute.Total.Acceleration.Load.EWMA=mean(Acute.Total.Acceleration.Load.EWMA))

filtered_Acwr <- na.omit(filtered_Acwr)
#optional cuts down 33k obs. to 6k obs.
```
```{r}
min(filtered_Acwr$Date,na.rm=TRUE)
max(filtered_Acwr$Date,na.rm=TRUE)
```

2. Catapult Session- Outdoor 
```{r}
catapult <- read.csv("~/T2P1-data-analysis-2025/data-sets/Catapult Session - Outdoor FB.csv")
```

```{r}
#remove periods + venues 
catapult <- catapult[,-c(16,19,20,21)]

#sort out any non football values
catapult <- catapult %>%
filter(Sport=="FOOTBALL")
catapult <- catapult[,-c(6)]  #remove sport column b/c not needed anymore

#cut back to last 6 months (can repeat to do last 3 months aswell or instead)
catapult$Date <- mdy(catapult$Date)
current_date <- Sys.Date()
start_date <- current_date- months(24)
catapult <- catapult %>%
  filter(Date >= start_date) %>%
  group_by(anon_id,Date, Primary.Position, Month, Year) %>%
  summarise(across(where(is.numeric), mean, na.rm=FALSE))
```
```{r}
filtered_catapult <- subset(catapult,select=c("anon_id","Date","Primary.Position","Position","Total.Distance","Maximum.Velocity","Average.Velocity","TRIMP", "Velocity.Exertion", "Total.Player.Load", "Player.Load.min","Total.Acceleration.Load","Acceleration.Density.Index","Explosive.Efforts","Max.Acceleration","Max.Deceleration","Session.Mean.TRIMP","HSD.SD","VHSD.SD","Typical.Player.Load","Typical.Player.Load","Typical.Player.Load.min","Average.IMA","Session.Classification"))

#can also add averaged bands(1-8) columns of diffrent metrics 
```
```{r}
min(filtered_catapult$Date,na.rm=TRUE)
max(filtered_catapult$Date,na.rm=TRUE)
```

##Just to look into other metrics, averaged (below)
```{r}
#combine date (3rd column) + athlete (2nd column // anon_id) to make an average/max or sum depending 
#group by (anon_id+Date), take the mean of all numerical columns 
Catapult <- catapult %>%
  group_by(anon_id, Date) %>%
  summarise(across(where(is.numeric), mean, na.rm=FALSE))
#Catapult dataset only has numerical values 

#Catapult<- na.omit(Catapult)
```

```{r}
#IMAs are counted variable so sum these
mean(catapult$High.Speed.Distance,na.rm=TRUE)
mean(catapult$Maximum.Velocity,na.rm=TRUE)
mean(catapult$Total.Player.Load,na.rm=TRUE)
mean(catapult$Player.Load.min,na.rm=TRUE)

#Maximum + average of the maximum & average velocity (given to us)
max(catapult$Maximum.Velocity,na.rm=TRUE)
mean(catapult$Average.Velocity,na.rm=TRUE)

#velocity + acceleration + deceleration + player load + IMA most important to pay attention to 
```

3. Incident Report
```{r}
incident <- read.csv("~/T2P1-data-analysis-2025/data-sets/Incident Report FB.csv")
```
```{r}
#sort out any non football values + etc
incident <- incident %>%
filter(Sport=="FOOTBALL") %>%
filter(Incident.Type=="Injury") %>%
filter(Result.of.Sport.Participation=="Yes")
#selecting the variables of interest within the incident report
filtered_incident <- subset(incident,select=c("anon_id", "Position", "Date.of.Injury...Onset.of.symptoms", "Side", "Body.Part", "Tissue.Type", "Pathology.Type", "OSICS14.Code", "OSICS10.Code", "OSICS10.Diagnosis", "Final.Diagnosis", "Result.of.Sport.Participation", "Recurrence.of.Injury", "Injury.Prognosis", "General.Mechanism", "Specific.Mechanism", "FinalDiagnosisIncidentDate", "Total.Time.Injured", "Month", "Year","Incident.Type"))

#cut back to last 6 months (can repeat to do last 3 months aswell or instead)
filtered_incident$Date <- mdy(filtered_incident$Date)
current_date <- Sys.Date()
start_date <- current_date- months(24)
filtered_incident <- filtered_incident %>%
  filter(Date >= start_date)

#HAMSTRING INJURY BINARY TO USE AS RESPONSE VARIABLE 
filtered_incident$HSI <- ifelse(filtered_incident$Body.Part == "Thigh",1,0)
```
```{r}
min(filtered_incident$Date,na.rm=TRUE)
max(filtered_incident$Date,na.rm=TRUE)
```

4. Performance Normative Data
```{r}
perfnorm <- read.csv("~/T2P1-data-analysis-2025/data-sets/Performance Normative Data FB.csv")
```
```{r}
filtered_perfnorm <- subset(perfnorm,select=c("anon_id","Date","Position","Nordic.Average","Nordic.Imbalance..","Adduction.Abduction.Ratio","Eccentric.Peak.Power...Relative","Jump.Height","Modified.RSI"))

#cut back to last 6 months (can repeat to do last 3 months aswell or instead)
filtered_perfnorm$Date <- mdy(filtered_perfnorm$Date)
current_date <- Sys.Date()
start_date <- current_date- months(24)
filtered_perfnorm <- filtered_perfnorm %>%
  filter(Date >= start_date) %>%
  group_by(anon_id, Date)

#can compare to NFL + NBA aswell

min(filtered_perfnorm$Date,na.rm=TRUE)
max(filtered_perfnorm$Date,na.rm=TRUE)
```

5.Performance Risk Assessment
##DATASET ONLY FROM ONE DAY OF LAST YEAR// THEREFORE PROBABLY NOT THAT RELEVANT TO US
#other than for possible comparisons 
```{r}
perfrisk <- read.csv("~/T2P1-data-analysis-2025/data-sets/Performance Risk Assessment FB.csv")

#filter football
perfrisk <- perfrisk %>%
  filter(Sport=="FOOTBALL")
```

```{r}
#ALL variables are relevant
filtered_perfrisk <- subset(perfrisk,select=c("anon_id", "Date","Position","Hamstring.Avg..Imbalance", "L..Hamstring.Max.Force","R..Hamstring.Max.Force","L..Eccentric.Mean.Force","R..Eccentric.Mean.Force","Eccentric.Mean.Force.Asym","CMJ.Landing.Asym.","CMJ.Jump.Height","Groin.Avg..Imbalance", "Relative.Peak.Power","CMJ.Peak.Force","IMTP.Peak.Force"))
```

```{r}
min(filtered_perfrisk$Date,na.rm=TRUE)
max(filtered_perfrisk$Date,na.rm=TRUE)
```

6. Soft Tissue Risk Status 
```{r}
riskstatus <- read.csv("~/T2P1-data-analysis-2025/data-sets/Soft Tissue Risk Status Report FB.csv")
```

```{r}
#refined risk status dataset
#contains response variable, what we are comparing against// validating 
riskstatus <- riskstatus[,-c(1,5)]

#cut back to last 6 months (can repeat to do last 3 months aswell or instead)
riskstatus$Date <- mdy(riskstatus$Date)
current_date <- Sys.Date()
start_date <- current_date- months(24)
riskstatus <- riskstatus %>%
  filter(Date >= start_date)  # 215 -> 100 obs.

#binary could be useful when testing accuracy/importance of high risk status
Highrisk_binary=ifelse(riskstatus$Status=="High",1,0)
#cbind(Riskstatus,Highrisk_binary)

riskstatus$Highrisk_binary = Highrisk_binary

min(riskstatus$Date,na.rm=TRUE)
max(riskstatus$Date,na.rm=TRUE)
```

7. Strength Testing Data
```{r}
#not expected to use this dataset alot + no real need to filter atm 
strengthtest <- read.csv("~/T2P1-data-analysis-2025/data-sets/Strength Testing Data FB.csv")
```

```{r}
#cut back to last 6 months (can repeat to do last 3 months aswell or instead)
strengthtest$Date <- mdy(strengthtest$Date)
current_date <- Sys.Date()
start_date <- current_date- months(24)
strengthtest <- strengthtest %>%
  filter(Date >= start_date)  # 224 -> 26 obs.

min(strengthtest$Date,na.rm=TRUE)
max(strengthtest$Date,na.rm=TRUE)
```

8. VALD Performance Test 
```{r}
VALD <- read.csv("~/T2P1-data-analysis-2025/data-sets/VALD - Performance Test FB.csv")
```
```{r}
#focus on NORDIC #s mainly (force,power, asymmetry, difference from normative to first) 

filtered_VALD <- subset(VALD, select = c("anon_id","Date", "Session.Date", "Position", "Trend", "Maximum.Force","Average.Force","Impulse", "Percent.Difference.Norm", "Nordic.Left.MAX", "Nordic.Right.MAX", "Nordic.MEAN.Imbalance.with.side", "Positional.Norm","Hip.Abd.Trend","Hip.Add.Trend", "Nordic...Difference.from.First.Test", "Month", "Year"))

#cut back to last 6 months (can repeat to do last 3 months aswell or instead)
filtered_VALD$Date <- mdy(filtered_VALD$Date)
current_date <- Sys.Date()
start_date <- current_date- months(24)
filtered_VALD <- filtered_VALD %>%
  filter(Date >= start_date)

min(filtered_VALD$Date,na.rm=TRUE)
max(filtered_VALD$Date,na.rm=TRUE)
```

## CARETERS
```{r}
#selecting variables of interest
Vald <- subset(VALD, select = c("anon_id", "Session.Date", "Position", "Trend", "Maximum.Force", "Percent.Difference.Norm", "Nordic.Left.MAX", "Nordic.Right.MAX", "Nordic.MEAN.Imbalance.with.side", "Positional.Norm", "Nordic...Difference.from.First.Test", "Month", "Year"))

vald <- Vald %>%
  mutate(Session.Date = as.Date(Session.Date, format = '%m/%d/%Y')) %>%
  group_by(anon_id, Month, Year, Trend) %>%
  summarise(Maximum.Force = mean(Maximum.Force, na.rm = TRUE),
            Percent.Difference.Norm = mean(Percent.Difference.Norm, na.rm = TRUE),
            Nordic.Left.MAX = mean(Nordic.Left.MAX, na.rm = TRUE),
            Nordic.Right.MAX = mean(Nordic.Right.MAX, na.rm = TRUE),
            Nordic.MEAN.Imbalance.with.side = mean(Nordic.MEAN.Imbalance.with.side, na.rm = TRUE),
            Positional.Norm = mean(Positional.Norm, na.rm = TRUE),
            Nordic...Difference.from.First.Test = mean(Nordic...Difference.from.First.Test, na.rm = TRUE)) %>%
  ungroup()
```

9. VALD Normative Report 
```{r}
#ALL variables relevant 
```

10. Wellness Report 
```{r}
wellness <- read.csv("~/T2P1-data-analysis-2025/data-sets/Wellness Report FB.csv")
```

```{r}
wellness <- wellness %>%
  filter(Sport=="FOOTBALL")

#mental score, physical score,sleep quality score, soreness scores, + overall wellness all relevant 
filtered_wellness <- subset(wellness, select=c("anon_id","Date","Position","Mental.Score","Physical.Score", "Sleep.Quality.Score","Soreness.Score","Overall.Wellness.Score","Wellness.Compliance"))

#cut back to last 6 months (can repeat to do last 3 months aswell or instead)
filtered_wellness$Date <- mdy(filtered_wellness$Date)
current_date <- Sys.Date()
start_date <- current_date- months(24)
filtered_wellness <- filtered_wellness %>%
  filter(Date >= start_date) 

min(filtered_wellness$Date,na.rm=TRUE)
max(filtered_wellness$Date,na.rm=TRUE)
```

## Wellness Dataset//CARTER
```{r}
#selecting only the variables of interest: important to not that this only contains data from may to december of 2024
Wellness <- subset(wellness, select = c("anon_id", "Date", "Mental.Score", "Physical.Score", "Sleep.Quality.Score", "Soreness.Score", "Overall.Wellness.Score", "Month", "Year"))

#omit the N/A's
Wellness2 <- Wellness %>%
  na.omit() %>%
  group_by(anon_id, Month, Year) %>%
  summarise(Mental.Score = mean(Mental.Score),
            Physical.Score = mean(Physical.Score),
            Sleep.Quality.Score = mean(Sleep.Quality.Score),
            Soreness.Score = mean(Soreness.Score),
            Overall.Wellness.Score = mean(Overall.Wellness.Score))
```


## 3 Months Prior Emphasis from Blake


```{r, warning = FALSE}
#generates a date for 3 months from the injury and then collapses the injury dates and 3 month prior dates into 8 columns (4 injuries per player) for each player 
HSI_3_months <- incident_HSI %>%
  mutate(Date = as.Date(Date.of.Injury...Onset.of.symptoms, format = '%m/%d/%Y'),
         start_date = Date - months(3)) %>%
  arrange(Date) %>%
  group_by(anon_id) %>%
  summarize(Dates.Final = paste0(Date, collapse = " and "),
            Dates.Start = paste0(start_date, collapse = " and "),
            Mechanism = paste0(General.Mechanism, collapse = " and "),
            Recurrence = paste0(Recurrence.of.Injury, collapse = " and "),
            Time.Injured = paste0(Total.Time.Injured, collapse = " and ")) %>%
  separate(Dates.Final, c('Injury.1', 'Injury.2', 'Injury.3', 'Injury.4'), sep = ' and ') %>%
  separate(Dates.Start, c('Injury.1.Start', 'Injury.2.Start', 'Injury.3.Start', 'Injury.4.Start'), sep = ' and ') %>%
  separate(Mechanism, c('Injury.1.Mech', 'Injury.2.Mech', 'Injury.3.Mech', 'Injury.4.Mech'), sep = ' and ') %>%
  separate(Recurrence, c('Injury.1.Recur', 'Injury.2.Recur', 'Injury.3.Recur', 'Injury.4.Recur'), sep = ' and ') %>%
  separate(Time.Injured, c('Injury.1.Time', 'Injury.2.Time', 'Injury.3.Time', 'Injury.4.Time'), sep = ' and ')
```

2. Merging VALD

```{r}
merge_vald <- merge(HSI_3_months, vald_merge, by = c('anon_id'), all = TRUE)

merge_vald <- merge_vald %>%
  mutate(Group = ifelse(Injury.1.Start <= Date & Date <= Injury.1, 1, ifelse(Injury.2.Start <= Date & Date <= Injury.2, 2, ifelse(Injury.3.Start <= Date & Date <= Injury.3, 3, ifelse(Injury.4.Start <= Date & Date <= Injury.4, 4, ifelse(Date > Injury.4, 'After', 0)))))) %>%
  filter(is.na(Group) | Group != 'After')

merge_vald <- merge_vald %>%
  group_by(anon_id, Group, Sport.Position) %>%
  summarise(Maximum.Force = mean(Maximum.Force, na.rm = TRUE),
            Percent.Difference.Norm = mean(Percent.Difference.Norm, na.rm = TRUE),
            Nordic.Left.MAX = mean(Nordic.Left.MAX, na.rm = TRUE),
            Nordic.Right.MAX = mean(Nordic.Right.MAX, na.rm = TRUE),
            Nordic.MEAN.Imbalance.with.side = mean(Nordic.MEAN.Imbalance.with.side, na.rm = TRUE),
            Positional.Norm = mean(Positional.Norm, na.rm = TRUE),
            Nordic...Difference.from.First.Test = mean(Nordic...Difference.from.First.Test, na.rm = TRUE))
```

```{r}
#trying to merge hamstring injuries with other lower body injuries
merge_lower <- merge(HSI_3_months, incident_lower, by = c('anon_id'), all = TRUE)

merge_lower <- merge_lower %>%
  mutate(Group = ifelse(Injury.1.Start <= Date & Date <= Injury.1, 1, ifelse(Injury.2.Start <= Date & Date <= Injury.2, 2, ifelse(Injury.3.Start <= Date & Date <= Injury.3, 3, ifelse(Injury.4.Start <= Date & Date <= Injury.4, 4, ifelse(Date > Injury.4, 'After', 0)))))) %>%
  filter(is.na(Group) | Group != 'After')

merge_lower <- merge_lower %>%
  mutate(Mechanism = ifelse(Group == 4, Injury.4.Mech, ifelse(Group == 3, Injury.3.Mech, ifelse(Group == 2, Injury.2.Mech, ifelse(Group == 1, Injury.1.Mech, NA)))),
         Recurrence = ifelse(Group == 4, Injury.4.Recur, ifelse(Group == 3, Injury.3.Recur, ifelse(Group == 2, Injury.2.Recur, ifelse(Group == 1, Injury.1.Recur, NA)))),
         Time.Injured = ifelse(Group == 4, Injury.4.Time, ifelse(Group == 3, Injury.3.Time, ifelse(Group == 2, Injury.2.Time, ifelse(Group == 1, Injury.1.Time, NA))))) %>%
  group_by(anon_id, Group, Mechanism, Recurrence, Time.Injured) %>%
  summarise(OSICS14.Code = paste0(OSICS14.Code, collapse = " and ")) %>%
  mutate(Injuries.Last.3.Months = ifelse(is.na(Group) | Group == 0, sapply(strsplit(OSICS14.Code, " and "), length), sapply(strsplit(OSICS14.Code, " and "), length) - 1))
```

```{r}
full_merge <- merge(merge_lower, merge_vald, by = c('anon_id', 'Group'), all = TRUE)

filtered_merge <- full_merge %>%
  mutate(Group = ifelse(is.na(Group), 0, Group),
         Mechanism = ifelse(is.na(Mechanism), 'None', Mechanism),
         Recurrence = ifelse(is.na(Recurrence), 'None', Recurrence),
         Time.Injured = ifelse(is.na(Time.Injured), 0, Time.Injured),
         OSICS14.Code = ifelse(is.na(OSICS14.Code), 'None', OSICS14.Code),
         HSI = ifelse(Group > 0, 1, 0)) %>%
  select(-Group) %>%
  na.omit()
```


## Validating the Soft-Tissue Watchlist

```{r}
#filtering just the high risk athletes from the watchlist
Riskstatus_high <- Riskstatus %>%
  filter(Status == 'High') %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  arrange(Date) %>%
  group_by(anon_id) %>%
  summarize(DateRisk = paste0(Date, collapse = " and "))

#creating a list of the unique athlete ids that were deemed high risk
high_risk <- unique(Riskstatus_high$anon_id)
```

```{r}
Incident <- read.csv("~/T2P1-data-analysis-2025/data-sets/Incident Report FB.csv")

#creating a list of the lower body areas in Body.Part and for OSICS codes for HSI
lower_body <- c("Thigh", "Ankle", "Lower Leg", "Knee", "Foot", "Groin/hip")
hamstring_OSICS14 <- c('TM1', 'TMB', 'TMS', 'TR1')

#filtering the incident report to only show those deemed high risk with only lower body injuries as well as only HSI related injuries
incident_hr_all <- Incident %>%
  filter(anon_id %in% high_risk, Body.Part %in% lower_body) %>%
  group_by(anon_id, Date.of.Injury...Onset.of.symptoms) %>%
  mutate(Total.Time.Injured = sum(Total.Time.Injured, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(across(everything())) %>%
  summarise(.groups = 'drop')

incident_hr_HSI <- Incident %>%
  filter(anon_id %in% high_risk, OSICS14.Code %in% hamstring_OSICS14) %>%
  group_by(anon_id, Date.of.Injury...Onset.of.symptoms) %>%
  mutate(Total.Time.Injured = sum(Total.Time.Injured, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(across(everything())) %>%
  summarise(.groups = 'drop')

#consolidates all injuries for these athletes into one column
incident_hr_all2 <- incident_hr_all %>%
  mutate(Date = as.Date(Date.of.Injury...Onset.of.symptoms, format = '%m/%d/%Y')) %>%
  arrange(Date) %>%
  group_by(anon_id) %>%
  summarize(Outcome = paste0(FinalDiagnosisIncidentDate, collapse = " and "))

incident_hr_HSI2 <- incident_hr_HSI %>%
  mutate(Date = as.Date(Date.of.Injury...Onset.of.symptoms, format = '%m/%d/%Y')) %>%
  arrange(Date) %>%
  group_by(anon_id) %>%
  summarize(Outcome = paste0(FinalDiagnosisIncidentDate, collapse = " and "))
```

```{r, warning = FALSE}
#combines the dates the athletes were deemed high risk with their injuries and separates the injury occurrences and dates
validation_hr_all <- left_join(Riskstatus_high, incident_hr_all2, by = 'anon_id') %>%
  separate(Outcome, c('Injury.1', 'Injury.2', 'Injury.3', 'Injury.4', 'Injury.5', 'Injury.6', 'Injury.7', 'Injury.8'), sep = ' and ') %>%
  separate(DateRisk, c('Test.1', 'Test.2'), sep = ' and ')

#combines the above but specifically for HSI and includes T/F for whether the athlete suffered their injury after their high risk assessment
validation_hr_HSI <- left_join(Riskstatus_high, incident_hr_HSI2, by = 'anon_id') %>%
  separate(Outcome, c('Injury.1', 'Injury.2', 'Injury.3', 'Injury.4'), sep = ' and ') %>%
  separate(DateRisk, c('Test.1', 'Test.2'), sep = ' and ')

validation_hr_HSI <- validation_hr_HSI %>%
  mutate(Injury.1.After = as.Date(str_sub(validation_hr_HSI$Injury.1, start = -10), format = '%m/%d/%Y') - as.Date(Test.1),
         Injury.2.After = as.Date(str_sub(validation_hr_HSI$Injury.2, start = -10), format = '%m/%d/%Y') - as.Date(Test.1),
         Injury.3.After = as.Date(str_sub(validation_hr_HSI$Injury.3, start = -10), format = '%m/%d/%Y') - as.Date(Test.1),
         Injury.4.After = as.Date(str_sub(validation_hr_HSI$Injury.4, start = -10), format = '%m/%d/%Y') - as.Date(Test.1),
         Injury.1.After.2 = as.Date(str_sub(validation_hr_HSI$Injury.1, start = -10), format = '%m/%d/%Y') - as.Date(Test.2),
         Injury.2.After.2 = as.Date(str_sub(validation_hr_HSI$Injury.2, start = -10), format = '%m/%d/%Y') - as.Date(Test.2),
         Injury.3.After.2 = as.Date(str_sub(validation_hr_HSI$Injury.3, start = -10), format = '%m/%d/%Y') - as.Date(Test.2),
         Injury.4.After.2 = as.Date(str_sub(validation_hr_HSI$Injury.4, start = -10), format = '%m/%d/%Y') - as.Date(Test.2))
         
#adds columns that return 'Yes' if an injury occurred after the athlete was deemed high risk and a 'No' if no injury occurred after
validation_hr_HSI <- validation_hr_HSI %>%
         mutate(Test.1.Success = ifelse(is.na(Test.1), NA, ifelse(rowSums((select(., Injury.1.After:Injury.4.After) > 0), na.rm = TRUE) > 0, 'Yes', 'No')),
         Test.2.Success = ifelse(is.na(Test.2), NA, ifelse(rowSums((select(., Injury.1.After.2:Injury.4.After.2) > 0), na.rm = TRUE) > 0, 'Yes', 'No')))
```

```{r}
#combining tests and injury outcomes and creating a data frame that just contains the dates of the test results and whether an injury occurred for the athlete after these tests
tests_hr <- c(validation_hr_HSI$Test.1, validation_hr_HSI$Test.2)
outcomes_hr <- c(validation_hr_HSI$Test.1.Success, validation_hr_HSI$Test.2.Success)
tests_outcomes_hr <- data.frame(tests_hr, outcomes_hr) %>%
  na.omit

#calculating percent of the time right
mean(tests_outcomes_hr$outcomes == 'Yes')

#visualizing using a bar plot 
ggplot(tests_outcomes_hr, aes(x = outcomes_hr, fill = outcomes_hr)) + 
  geom_bar(color = 'black', show.legend = FALSE) +
  scale_fill_manual(values = c('gold2', 'grey2')) + 
  labs(title = 'Outcomes of Soft Tissue Watchlist (High Risk)',
       x = 'Injury Occured After Deemed High Risk',
       y = 'Number of Observations') + 
  theme_bw()
```

```{r}
#important date of injury, injury type, +OSICIS 10 (already standaridzed codification of injury type), time loss, recurring injury 
#rest irrelevant// get rid of 
Incident <- subset(incident,select=c("anon_id","Date","Position","Date.of.Injury...Onset.of.symptoms","Side","Body.Part","OSICS.Injury.Diagnosis","Coach.s.Diagnosis","OSICS10.Diagnosis","Recurrence.of.Injury","Injury.Prognosis","Total.Time.Injured"))
```


```{r}
#MERGE Incident and Riskstatus datasets
df <- merge(riskstatus,filtered_incident,by="anon_id")  #26 predictors 
```
```{r}
df <- merge(df,filtered_perfnorm,by="anon_id")  #34 predictors
```
```{r}
df <- merge(df,filtered_wellness,by="anon_id")
```

## Exploratory Analysis: descriptive stats + plots
```{r}
#density plot of risk categories 
ggplot(df, aes(x=Status))+
  geom_bar(fill="lightblue")+
  labs(title="Risk Status Spread",x="Risk Status",y="Frequency")

ggplot(df, aes(x=Position))+
  geom_bar(fill="lightblue")+
  labs(title="Positional Spread",x="Players Positions",y="Frequency")

#position by status
ggplot(df,aes(x=Position,y=Status))+  #scatterplot
  geom_point()+
  geom_jitter()+ 
  labs(title="Position Influence on Hamstring Risk Status",y="Risk Status")

ggplot(df,aes(x=Position,fill=Status))+   #stacked bar chart
  geom_bar(inherit.aes = TRUE)+
  scale_fill_brewer()
```
```{r}
#density plot of Body. part density 
ggplot(df, aes(x=Body.Part))+
  geom_bar(fill="lightblue")+
  labs(title="Part of Body INjured Spread",x="Injured Body Part",y="Frequency")

ggplot(df, aes(x=Recurrence.of.Injury))+
  geom_bar(fill="lightblue")+
  labs(title="Recurrence of Injury Spread",x="Injury Recurrence",y="Frequency")


#injury rate by risk category (Riskstatus)
#injury rate by previous injury history

#distributions of strength/imbalance by risk category 
#boxplot(y~`Status,data=df)
```


```{r}
#risk category validity checks 
```

```{r}
#correlation checks 
```

```{r}